% - http://blog.revolutionanalytics.com/2009/09/the-difference-between-statistics-and-machine-learning.html

% - http://stats.stackexchange.com/questions/6/the-two-cultures-statistics-vs-machine-learning



Dendrograms and hierarchical clustering.
K nearest neighbour analysis.
Gibbs Sampling
Metropolis Hastings Algorithm
Jacard Distance
Bray Curtis Disimilarityo
Euclidean Distance
Mahalanobis Distance covariance matrix.
Clique. Used in Sodial Network Analysis. Degree.
Coplots and biplots.
Sorensen Distance
vegan r package
Principal Component Analysis
Data Reduction Technique
Multicollinearity
Hadoop Pig and Mahout
Beer Lambert Law for removing the intercept
Markov Random Field is an n dimensional random process defined on a discrete lattice. Usually the lattice is a regular 2 dimensional in the plane , either finite or infinite . The conditional distribution of Xn depends  only on Xn-1 and Xn+1.
Spurious Correlation
Time to Event Data in Survival analysis  
Ising Model
Nested and orthogonal effects
Hopfield Networks
Form of recurrent artificial neural networks that serve as content addressable memory systems with binary threshold unitts. There is an underlying Luapuniv function for the activity dynamics.
 2 popular forms of the model
 binary neurons with discrete time updated one at a time.
Graded neuron with continuous time
activity at neuron j
mean ibternal potential at neuron j
direct input to the neuron 
strength of synaptic input from k to j.
Gibbs distribution
 P (x)  
Z is a normalizing constant called the partition function.T is a constant called temperature. U is the energy function.
