% - MODULE 5: Further probability and inference

\subsection*{Bivariate distributions}

Simple bivariate discrete distributions. 
Joint, conditional and marginal distributions: probability mass function, expectation and variance. 
Covariance and correlation.

Distributions with probabilities presented either by formula or in a two-way table.
Simple bivariate continuous distributions. 


Joint, conditional and marginal distributions: probability density function, expectation and variance. 
Covariance and correlation.

Simple cases only – rectangular and triangular spaces, with one or more sides parallel to a coordinate axis.
The bivariate Normal distribution.
%------------------------------------------------------------------------------%
\textit{Familiarity with the bivariate Normal distribution as a model (knowledge of joint pdf is not required). 
Proofs of results are not required.}
%------------------------------------------------------------------------------%
\subsection*{Generating functions}
Probability and moment generating functions.
Use to find expectations and variances. 
Use to establish the distribution of sums of random variables.
No limiting results.
%------------------------------------------------------------------------------%
\subsection*{Inference}
\begin{itemize}
\item	The likelihood function.
\item	Plots of likelihood against the parameter.
\item	Estimation of a single parameter of a distribution using the method of moments and the method of maximum likelihood.
\item	Maximisation of likelihood using graphical methods and calculus. Questions may be set involving standard distributions listed in Module 2 and other simple cases.
\item	Properties of point estimators. Range, unbiasedness, consistency. Efficiency and relative efficiency.
\item	Introductory treatment of these concepts in simple cases.
\item	Calculation of approximate variance of a maximum likelihood estimator using second derivative of log likelihood
\end{itemize}
%------------------------------------------------------------------------------%
